### 2021-02-19 architecture.tex
Remove some redundant content from Architecture
Name a section with alternative content
Add dummy subsections in Architecture
zatim par poznamek

The architecture is designed so that the recommender can be trained
on the results executions of an \gls{atp} on various problems with random precedences.
The recommender is parameterized by weight tensors
whose values can be trained by gradient descent.
\todo{Ensure we have defined ``gradient descent''.}
The goal of the training is to find parameter values
such that the recommender produces precedences
that are likely to yield a successful proof in few iterations of the saturation loop.
\todo{FB: Too vague?}

The recommender consists of modules that perform specific subtasks (see \cref{fig:architecture}):
\begin{enumerate}
\item Graphifier converts the input \gls{cnf} problem into a problem graph.
\item \gls{gcn} converts a problem graph into symbol embeddings.
The \gls{gcn} contains one or more convolution layers.
\item Projection converts the symbol embeddings into a vector of symbol costs.
\item Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{enumerate}

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous directed graph}.\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
\todo{FB: Consider removing the footnote because it is basically useless.}
Each of the nodes is labeled with a node type,
and each edge is labeled with an edge type,
informing the heterogeneous nature of the graph.
The nodes of the graph correspond to various elements that constitute the \gls{cnf} formula,
such as clauses, atoms and predicate symbols.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the relation between a clause and one of its atoms,
or the relation between an atom and its predicate symbol.
\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.

@@

The graph representation has the following properties:
\begin{itemize}
\item For each relation $r$, its inverse $\inv{r}$ is also present in the graph,
represented by a different edge type (unless $r = \inv{r}$ is an identity).
% Note: The relation is considered an identity if it is a subset of the complete identity relation.
\item Lossless: The original problem can be reconstructed from the graph representation completely
(up to equivalent modifications).
\item Signature agnostic: Renaming the symbols and variables in the input problem yields an isomorphic graph.
\item A singleton node of type \ntype{problem} is connected to all the clauses of the problem.
\item The polarity of the literals is expressed by the type of the edge (\epos{} or \eneg{})
connecting the respective atom to the clause it occurs in.
\item For every non-equality atom and term, the order of its arguments is captured by a sequence of \ntype{argument} nodes chained by edges \cite{Rawson2020}.
\item The two operands of equality are not ordered.
This reflects the symmetry of equality.
\item Term-sharing is employed: Redundant atoms and terms share a node representation \cite{}.
Note that since each variable is bound by a clause,
ground terms may be shared across clauses,
but non-ground terms may only be shared within a clause.
\todo{FB: Remove as of little interest?}
\end{itemize}
\todo{FB: Reference appendix if we describe the representation in more detail there.}
\todo[inline]{MS: Filosoficka k ``can be represented''; mozna by se mohlo nekde objevit:
Snazime se do struktury HIN ``otisknout'' celou strukturu CNF (jak jsem rikal,
znamena to, ze zobrazeni je proste), abychom siti nevzali moznost ``vedet o problemu
uplne vsechno'' (az na veci, na kterych nezalezi jako poradi literaly v klauzulich, etc).
Ta filosoficka poznamka je o tom, ze pro dobre uceni tohle vlastne vubec nemusi byt treba,
ze nejaky ``hrubsi otisk'' by mohl stacit tez (a byt treba mensi),
ale my nevime, jak takovy zvolit. }

@@

For each symbol in the problem,
we seek to find a vector representation, i.e., an \emph{embedding},
that captures the symbol's properties that are relevant
for the desirable relative rank of the symbol in the symbol precedence.
The symbol embeddings are modeled by a \gls{rgcn} \cite{Schlichtkrull2017},
\todo{Can I use the term \gls{rgcn} when I use a modified version? Namely, the activation is called earlier.}
which is a stack of graph convolutional layers.
Each layer consists of a collection of differentiable modules --- one for each edge type.
A pass through the convolutional layer takes the embeddings of all symbols,
converts them by the edge modules of the adjacent edges to form the messages,
passes the messages to the destination nodes
and aggregates the incoming messages of each node.
In the simple case that each edge module is a dense layer,
the output embeddings $h^{(l+1)}$ is computed from the input embeddings $h^{(l)}$ in the following manner:
$$
h_i^{(l+1)} =
\sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in \mathcal{N}_i^r} \frac{1}{c_{ji}} h_j^{(l)} W_r^{(l)}}
$$
%$$
%h_i^{(l+1)} =
%\mathrm{LayerNorm} \Parentheses{h_i^{(l)} + \sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in %\mathcal{N}_i^r} \frac{1}{c_{ji}} h_j^{(l)} W_r^{(l)}}}
%$$
% Inspiration: https://ufal.mff.cuni.cz/~straka/courses/npfl114/1920/slides.pdf/npfl114-07.pdf - slide 27 - Transformer
\todo{FB: Mention inspiration: Transformer. See Straka's slides.}
$h_i^{(l)}$ denotes the embedding of node $i$ at layer $l$.
$\mathcal{R}$ denotes the set of all relations in the graph.
$\sigma$ denotes the activation function, for example the \gls{relu}.
$\mathcal{N}_i^r$ denotes the set of neighbors of node $i$ under relation $r$.
$c_{j, i, r} = \sqrt{\card{\mathcal{N}_j^r}} \sqrt{\card{\mathcal{N}_i^r}}$ is a normalization constant \cite{kipf2017semisupervised}.
$W_r^{(l)}$ is a trainable weight matrix representing relation $r$ at layer $l$.

@@

@@

\subsection{Notes}

The recommender consists of the following components:

\begin{itemize}
\item Symbol embedding model
\item Symbol cost model
\item Loss function
\item Precedence generator
\end{itemize}

>>>>

The architecture is designed so that the recommender can be trained
on the results of \gls{atp} executions on various problems with random precedences.
The recommender is parameterized by weight tensors
whose values can be trained by gradient descent.
\todo{Ensure we have defined ``gradient descent''.}
\todo{MS: chapu ucel (a vidim, ze dalsi veta navazuje na ``parameters'', ale je to kostrbate.
Recommender = system, chapu jako program. Programy vetsinou neparametrizujeme tenzorama a gradient descent zije na jine urovni,
nez aby parametrizovane programy primo ucil. Uceni je spis nejaka technologie, kterou recommender pouziva, aby ... ``produces precedences
that are likely to yield a successful proof''.  }
The goal of the training is to find parameter values
such that the recommender produces precedences
that are likely to yield a successful proof in few iterations of the saturation loop.
\todo{FB: Too vague? MS: v tomhle momente mi to prijde dobry!}

The recommender consists of modules that perform specific subtasks (see \cref{fig:architecture}):
\begin{enumerate}
\item Graphifier\todo{MS: Zatim nevidim Graphifier ve \cref{fig:architecture}.} converts the input \gls{cnf} problem into a problem graph.
\item \gls{gcn} converts a problem graph into symbol embeddings.
The \gls{gcn} contains one or more convolution layers.
\item Projection converts the symbol embeddings into a vector of symbol costs.
\item Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{enumerate}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... }

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous directed graph}.\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
\todo{FB: Consider removing the footnote because it is basically useless.}
\todo{MS: Footnote asi remove, ale neexistuje citace pro ``heterogeneous directed graph''?}
Each of the nodes is labeled with a node type,
and each edge is labeled with an edge type,
informing\todo{MS: maybe better ``defining''?} the heterogeneous nature of the graph.
The nodes of the graph correspond to various elements that constitute the \gls{cnf} formula,
such as predicate symbols, atoms, and clauses.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the incidence relation between a clause and one of its (literal's) atoms,
or the relation between an atom and its predicate symbol.
\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.

@@

The graph representation has the following properties:
\begin{itemize}
\item For each relation $r$, its inverse $\inv{r}$ is also present in the graph,
represented by a different edge type (unless $r = \inv{r}$ is an identity).
% Note: The relation is considered an identity if it is a subset of the complete identity relation.
\todo{MS: tohle jsem nepochopil. Nemuzeme rict, ze (az na nejakou vyjimkou) jsou edges v obou smerech?}
\item Lossless: The original problem can be faithfully reconstructed from the corresponding graph representation
(up to logical equivalence).
\item Signature agnostic: Renaming the symbols and variables in the input problem yields an isomorphic graph.
\item A singleton node of type \ntype{problem} is connected to all the clauses of the problem.
\item The polarity of the literals is expressed by the type of the edge (\epos{} or \eneg{})
connecting the respective atom to the clause it occurs in.
\item For every non-equality atom and term, the order of its arguments is captured by a sequence of \ntype{argument} nodes chained by edges \cite{Rawson2020}.
\item The two operands of equality are not ordered.
This reflects the symmetry of equality.
\item Term-sharing is employed:\todo{MS: Mozna presneji ``perfect sub-expression sharing''? 
Predpokladam ze dva ruzne ground literaly p(c) tam budou taky jen jendnou?}
Redundant atoms and terms share a node representation \cite{}.
Note that since each variable is bound by a clause,
ground terms may be shared across clauses,
but non-ground terms may only be shared within a clause.
\todo{FB: Remove as of little interest?}
\end{itemize}
\todo{FB: Reference appendix if we describe the representation in more detail there.}
\todo[inline]{MS: Filosoficka k ``can be represented''; mozna by se mohlo nekde objevit:
Snazime se do struktury HIN ``otisknout'' celou strukturu CNF (jak jsem rikal,
znamena to, ze zobrazeni je proste), abychom siti nevzali moznost ``vedet o problemu
uplne vsechno'' (az na veci, na kterych nezalezi jako poradi literaly v klauzulich, etc).
Ta filosoficka poznamka je o tom, ze pro dobre uceni tohle vlastne vubec nemusi byt treba,
ze nejaky ``hrubsi otisk'' by mohl stacit tez (a byt treba mensi), ale my nevime, jak takovy zvolit. }

@@

For each symbol in the problem,
we seek to find a vector representation, i.e., an \emph{embedding},
that captures the symbol's properties that are relevant
for the correctly ranking the symbol in the observed symbol precedences.
The symbol embeddings are modeled\todo{MS: embeddings are ``output''? } by a \gls{rgcn} \cite{Schlichtkrull2017},
\todo{Can I use the term \gls{rgcn} when I use a modified version? Namely, the activation is called earlier. MS: sounds good.}
which is a stack of graph convolutional layers.
Each layer consists of a collection of differentiable modules --- one for each edge type.
A pass through the convolutional layer takes the embeddings of all symbols,
converts them by the edge modules of the adjacent edges to form the messages,\todo{MS: ``message'' occurs for the first time. 
Maybe ``a message'', or explain beforehand.}
passes the messages to the destination nodes
and aggregates the incoming messages of each node.
In the simple case that each edge module is a dense layer,\todo{MS: hnidopich nevi, co to znamena ``dense layer''}
the output embedding $h^{(l+1)}$ is computed from the input embeddings $h^{(l)}$ in the following manner:
$$
h_i^{(l+1)} =
\sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in \mathcal{N}_i^r} \frac{1}{c_{ji}} h_j^{(l)} W_r^{(l)}}
$$
%$$
%h_i^{(l+1)} =
%\mathrm{LayerNorm} \Parentheses{h_i^{(l)} + \sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in %\mathcal{N}_i^r} \frac{1}{c_{ji}} h_j^{(l)} W_r^{(l)}}}
%$$
% Inspiration: https://ufal.mff.cuni.cz/~straka/courses/npfl114/1920/slides.pdf/npfl114-07.pdf - slide 27 - Transformer
\todo{FB: Mention inspiration: Transformer. See Straka's slides.}
$h_i^{(l)}$ denotes the embedding of node $i$ at layer $l$.
$\mathcal{R}$ denotes the set of all relations in the graph.
$\sigma$ denotes the activation function, for example the \gls{relu}.
$\mathcal{N}_i^r$ denotes the set of neighbors of node $i$ under relation $r$.
$c_{j, i, r} = \sqrt{\card{\mathcal{N}_j^r}} \sqrt{\card{\mathcal{N}_i^r}}$ is a normalization constant \cite{kipf2017semisupervised}.
$W_r^{(l)}$ is a trainable weight matrix representing relation $r$ at layer $l$.

@@

\subsection{Projection: From symbol embeddings to symbol costs}

\subsection{Sort: From symbol costs to precedence}

\subsection{Loss}

\subsection{[Alternative: Top-down approach]}

@@
